{How do we copy and paste?\\The semantic drift of quotations in blogspace}

### 2015-02-10 brains-protocol.tex
corrections generales sur la section 4, avec commentaires [le texte peut encore etre simplifie/clarifie]

With the substitution detection model thus defined, we may now formulate a family of psycholinguistic hypotheses describing the role of each feature in the accuracy of the reformulation.  To this end, we build two main observables for each word feature.  
First, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution under this condition. Second, we measure the variation of word feature over a substitution, looking at the variation of a given feature between start and arrival words.

Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration ---~e.g. drafting before publishing, double-checking sources, proof-reading~---, and may not be representative of the cognitive processes at work during alteration.
The two situations are different enough to be studied separately, and we focus here on the latter.

@@

Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$% (only taking into account words that are substituted at least once)
:
\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
$$\sigma_{\phi}(f) = \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$

@@

All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to an effect or to no effect of the selected psycholinguistic features on substitution susceptibility.
We may therefore globally assume that substitution targets are chosen in a more or less uniform way with respect to these features.

@@

Of prime interest is the comparison of the value of $\nu_{\phi}(f)$ with respect to $f$, as it shows whether there is an attraction (or a repulsion) effect towards (resp. from) some values of each feature.
In other words, plotting the $y=x$ line, we can see if substitutions tend to converge towards some typical value of a word feature or not --- as is classically done in the study of dynamical systems.

We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the pool of free association words.
In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words. In this case $\warrival_{00}$ does depend on $\wstart$. The results were not qualitatively different under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.}
$$\nu_{\phi}^0 (f) = \left<\phi\right>$$
%We thereby obtain the mean variation of feature for each start feature value, comparing the variations to a situation where arrival words are chosen randomly.

This approach yields a fine-grained view of how word features evolve upon substitution, on average, with respect to (i) the original feature (\hbox{vs.} $y=x$) and (ii) a random arrival (\hbox{vs.} $\mathcal{H}_0$).
\begin{figure*}[!th]
    \centering
    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
    \caption{\textbf{Feature variation upon substitution:} average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.    
    }
    \label{fig:feature-variations}
\end{figure*}

\medskip
Results are gathered on Fig.~\ref{fig:feature-variations}.
We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1, independently of the feature considered.
In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibits a unique attractor.
Second, the comparison with $\mathcal{H}_0$ shows that there are two classes of attractors, depending on whether:
\begin{enumerate}[(i)] 
\item there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$);
\item or $\nu_{\phi}$ remains always above or below $\nu_{\phi}^0$.
\end{enumerate}
The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed.
%\TB{(furthermore, $\mathcal{H}_{00}$ stays inside the confidence intervals of $\nu_{\phi}$, confirming that these two features are irrelevant for the substitution process)}.
%\rk{Cam: On vire ou bien?}

On the other hand, the second class (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear bias, positive or negative.

@@

@@

\medskip
We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value (i.e. $\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age of 8, and a Clustering coefficient of 0.1).

>>>>

We may now use this substitution model to formulate a family of psycholinguistic hypotheses describing the role of each feature in the accuracy of the reformulation.  To this end, we build two main observables for each word feature.  
First, we measure the susceptibility for words to be the target of a substitution in a quote, knowing that there has been a variation, in order to show which semantic features are the most likely to ``attract'' a substitution under this condition. Second, we measure the evolution of word feature over a substitution, looking at the variation of a given feature between start and arrival words.

Note that since we only consider substitutions and not faithful copies, we measure the features of an alteration \emph{knowing that there has been an alteration}, and we do not take invariant quotations into account.
Indeed, in the former case we know there has been a human reformulation, whereas in the latter case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').
Furthermore, perfect human reformulation possibly involves different practices than those involved in alteration ---~e.g. drafting before publishing, double-checking sources, proof-reading~--- and may not be representative of the cognitive processes at work during alteration.
The two situations are different enough to be studied separately, and we focus here on the latter.

@@

Now, for a given feature $\phi$, we obtain the mean susceptibility $\sigma_{\phi}(f)$ for the feature value $f$ by averaging this ratio over all words such that $\phi(w) = f$, that is% (only taking into account words that are substituted at least once)
:
\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.}
$$\sigma_{\phi}(f) = \left< \frac{s_w}{p_w} \right>_{\left\lbrace w | \phi(w) = f \right\rbrace}$$

@@

All in all, apart from Word frequency and despite some tendencies these results do not allow us to conclude either to a marked effect %or to no effect 
of the selected psycholinguistic features on substitution susceptibility.
We may therefore globally assume that substitution targets are chosen in a more or less uniform way with respect to these features.

@@

Of prime interest is the comparison of the value of $\nu_{\phi}(f)$ with respect to $f$, as it shows whether there is an attraction (or a repulsion) effect towards (respectively from) some values of each feature.
In other words, plotting the $y=x$ line, we can see if substitutions tend to converge towards some typical value of a word feature or not --- as is classically done in the study of dynamical systems.

We also introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival$ has been chosen randomly from the whole pool of free association words.\marginnote{C: ici on est bien d'accord que le pool of FA words c'est tout le set de mots, pas juste les voisins? la formulation est un peu ambigüe}
In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore),  the baseline variation under $\mathcal{H}_0$ may be rewritten as:\footnote{We additionally considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is randomly chosen \emph{among immediate synonyms of the start word}, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words. In this case $\warrival_{00}$ does depend on $\wstart$. Our conclusions hold under this second null hypothesis, so for the sake of clarity we chose to keep the simpler~$\mathcal{H}_0$.}
$$\nu_{\phi}^0 (f) = \left<\phi\right>$$
%We thereby obtain the mean variation of feature for each start feature value, comparing the variations to a situation where arrival words are chosen randomly.

This approach yields a fine-grained view of how word features evolve upon substitution, on average, with respect to (i) the original feature (\hbox{vs.} $y=x$) and (ii) a random arrival (\hbox{vs.} $\mathcal{H}_0$).
\begin{figure*}[!th]
    \centering
    \includegraphics[width=\textwidth]{images/computed-figures/feature_variations-binned.png}
    \caption{\textbf{Feature variation upon substitution:} $\nu_\phi$, average feature value of the appearing word as a function of the feature value of the disappearing word in a substitution, with 95\% asymptotic confidence intervals.
    The overall position of the curve with respect to the dashed line representing $\mathcal{H}_0$ (constant $\nu_{\phi}^0$) indicates the direction of the cognitive bias.
    %The fact that all curves have slopes smaller than 1 means that the substitution operation is contractile on average: each feature will converge towards its own specific asymptotic range.    
    }
    \label{fig:feature-variations}
\end{figure*}

\medskip
Results are gathered on Fig.~\ref{fig:feature-variations}.
We can do a first striking observation: all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, while the slope of $\nu_{\phi}$ is smaller than 1, independently of the feature considered.
In other words, beyond individual variation patterns, the substitution process is contractile for all the features, and each of them therefore exhibits a unique attractor.
Second, the comparison with $\mathcal{H}_0$ shows that there are two classes of attractors, depending on whether:
\begin{enumerate}[(i)] 
\item there is a triple intersection (of $y=x$, $\nu_{\phi}^0$ and $\nu_{\phi}$);
\item or $\nu_{\phi}$ always remains above or below $\nu_{\phi}^0$.
\end{enumerate}
The first class (Number of phonemes and Number of synonyms) are features for which the substitution process only brings words slightly closer to $\nu_{\phi}^0$, and no uniform bias can be observed.
%\TB{(furthermore, $\mathcal{H}_{00}$ stays inside the confidence intervals of $\nu_{\phi}$, confirming that these two features are irrelevant for the substitution process)}.
%\rk{Cam: On vire ou bien?}

On the other hand, the second class (comprising Word frequency, Age of acquisition, and Clustering coefficient) are features for which the substitution process has a clear bias, positive or negative, with respect to the purely random situation ($\mathcal{H}_0$).

@@

\marginnote{C: les légendes sur la figure 5 sont un peu trompeuses, on parle de $\nu_\phi$, $\nu_{\phi}^0$ dans le texte et ici de $y=x$ et $\mathcal{H}_0$; peut être préciser que la courbe pleine est $\nu_\phi$, etc.?}

@@

\medskip
We thus observe a clear convergence pattern for each feature, with two different classes corresponding to the psychological relevance of each feature for the substitution process.
Taken as a dynamical system where substitutions are repeatedly applied, Number of phonemes and Number of synonyms will simply converge towards their average value in the FA corpus (i.e. $\nu_{\phi}^0)$, while Word frequency, Age of acquisition and Clustering coefficient, consistent with the literature, will converge towards significantly biased values indicated by the intersection with $y = x$ (respectively, a frequency of $\exp(9.1) \simeq 9000$, an acquisition age of 8, and a Clustering coefficient of 0.1).
